% mainfile: ../ltexpprt.tex
We used a number of different data sources from widely different domains
such as physical indicators like weather characterisitcs to social
networking/user queries. For a number of sources such as Twitter data,
the most important part is to properly identify the The keywords
associated with flu so that we can get a proper estimate of the flu
realted surrogate information. In section~\ref{sec:keyword}, we describe
our keyword collection mechanism as follows.
\subsection{\label{sec:keyword} Keywords Extraction}
\input{tex/keywords}

\subsection{Data sources}
We next desribe the different data sources as follows: 
\subsubsection{Twitter} 
\subsubsection{Healthmap}
\subsubsection{Weather Data}
\subsubsection{Google Flu Trends}
\subsubsection{Google Search Trends}
\subsubsection{Open table}
We monitored table availability using OpenTable; an online restaurant
table reservation site for cities in the USA and Mexico. Our analysis
can be summarized as follows. First, using the OpenTable site, we
searched for the number of restaurants with available tables for two
persons at lunch and dinner. Since different regions and individuals
have different eating habits, we defined the lunch period between
12–3:30pm and dinner between 6–10:30pm. We searched for available tables
every hour and half past the hour for every day of the week. Next, we
investigated any occurrences of social unrest and natural disasters,
which might have affected the trend in the time series. Lastly, using
moving averages, cross-correlations and regression models, we elucidated
and compared the time-trend in the data of table availabilities to data
collected for various disease outbreaks. In the USA, we examined table
availability for restaurants in Boston, Atlanta, Baltimore and Miami.
For Mexico, we studied table availabilities in Cancun, Mexico City,
Puebla, Monterrey, and Guadalajara.
--We started collecting data on table availabilities starting September
4th 2012.
--The number of restaurants available varies over time since restaurants
are added and removed from the site. At the time of this writing, there
were over 28,000 restaurants in the database.
--Since we monitored ten regions at twenty search times, this resulted
in 200 distinct time-series curves.
--Size of data: The data files are not too big. When I analyzed it, each
was about 1kb which implies about 200kb in total.


